---
title: "Data Science and Analytics Assignment"
author: "Luigui Gallardo-Becerra"
date: "4/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("ggpubr")
library("optparse")
library("ggplot2")
library("reshape")
require("gridExtra")
library("ggpubr")
```

## Statistical analysis

### Is there a statistically significant relationship between the cancellation policy and unit price?

First, we need to import our data (listings.csv) into R.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Data input (listings.csv)
listings <- read.csv("dataset/listings.csv",
                     header=TRUE)
```

Looking at the data, the "price" contains the "$" and "," symbols. It is easier to manage the data if we remove these character.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Remotion of $ and , characters
listings$price = gsub("\\$", "", listings$price)
listings$price = gsub(",", "", listings$price)
listings$price = as.numeric(listings$price)
```

Before evaluate the relationship between these variables, we must know the distribution of our data, and with it select a parametric or non-parametric method. To achieve this we can use the Shapiro-Wilk's test.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Shapiro-Wilk's test on Price
shapiro.test(listings$price)
```

From this result we can assume that the data is significantly different to the normal distribution (p-value < 0.05), so we must use non-parametric tests.

Now we can answer the question. We can use a Kruskal-Wallis to compare the means between groups.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Kruskal-Wallis to contrast between groups
kruskal.test(data = listings, price ~ cancellation_policy)
```
With this result we can conclude that there are differences in the price between the cancellation policy groups.

To know which groups differ from each other, we can use a Wilconxon's test. Additionally we can obtain a p-value adjustment with FDR (False Discovery Rate).

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Wilcoxon's to contrast between groups
pairwise.wilcox.test(listings$price,
                     listings$cancellation_policy,
                     p.adjust.method = "fdr")
```

The p-values obtained were much lower than 0.05, so with this analysis we can conclude that the three groups means differ from each other.

Additionally, we can obtain a boxplot to visualize these results.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# P-value groups
p_value <- list(c("flexible", "moderate"), 
                    c("flexible", "strict"),
                    c("moderate", "strict"))
# Colors
colors <- c("blue",
            "orange",
            "red")

theme_set(theme_bw())

# Boxplot cancellation policy vs unit price
boxplot <- ggplot(listings,
                  aes(x = cancellation_policy,
                      y = price)) +
  geom_boxplot(aes(color = cancellation_policy)) +
  labs(x="Cancellation Policy", 
       y="Price ($)") +
  theme(legend.position = "none") +
  scale_color_manual(values = colors) +
  stat_compare_means(comparisons = p_value,
                     method = "wilcox.test",
                     size = 2)

boxplot
```

\newpage

### Construct a linear model to evaluate how the price of a rental is influenced by unit size, the number of bedrooms, the number of bathrooms, and the maximum occupancy. What can you conclude about the influence of these attributes on price?

We can use the same data that the previous analysis. Now we can make a Linear Model considering these factors.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
linear_model <- lm(data=listings, price ~ square_feet + bedrooms + bathrooms + guests_included)

summary(linear_model)
```

With these results we can conclude than there is a significant association between the price and the unit size (0.03), maximum occupancy (0.0005), number of bedrooms (8.7e-10). The number of bathrooms was not significantly associated with the price.

Additionally, we can obtain plots to visualize these results.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
theme_set(theme_bw())
# Linear models of price vs square_feet + bedrooms + bathrooms + guests_included
# Square feet
lm_square_feet <- ggplot(listings,
                         aes(y = price,
                             x = square_feet)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(x="Square feet", 
       y="Price ($)") +
  theme(legend.position = "none")

# Bedrooms
lm_bedrooms <- ggplot(listings,
                      aes(y = price,
                          x = bedrooms)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(x="Number of bedrooms", 
       y="Price ($)") +
  theme(legend.position = "none")

# Bathrooms
lm_bathrooms <- ggplot(listings,
                       aes(y = price,
                           x = bathrooms)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(x="Number of bathrooms", 
       y="Price ($)") +
  theme(legend.position = "none")

# Guest included
lm_guests_included <- ggplot(listings,
                             aes(y = price,
                                 x = guests_included)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(x="Guest included", 
       y="Price ($)") +
  theme(legend.position = "none")

# Final Plot
ggarrange(lm_square_feet,
          lm_bedrooms,
          lm_bathrooms,
          lm_guests_included,
          ncol = 2, nrow = 2)
```
